---
title: "Class11"
author: "LVN"
date: "May 9, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB statistics
Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. From the website what proportion of structures are protein?

```{r}
PDBstatistics <- read.csv("Data Export Summary.csv", row.names = 1)
percent <- PDBstatistics$Total / sum(PDBstatistics$Total)
names(percent) <- row.names(PDBstatistics)
barplot(percent)
```


## Using Bio3D

```{r}
library(bio3d)
# in the console, type lbio3d to list all the functions in the package. help(package=bio3d) will open a help page for those functions.
pdb <- read.pdb("1hsg") ## Note: Accessing on-line PDB file
pdb
```

```{r}
attributes(pdb)
```


```{r}
head(pdb$atom)
```

```{r}
pdb$atom[1:2, c("eleno", "elety", "x","y","z")] # Print a subset of $atom data for the first two atoms
```


```{r}
# Note that individual $atom records can also be accessed like this
pdb$atom$elety[1:2]
```




```{r}
# Which allows us to do the following
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab="B-factor")
```



Q7. What type of R object is pdb$atom? HINT: You can always use the str() function to get a useful summery of any R object.
```{r}
str(pdb$atom) #pdt$atom is a data frame
```


```{r}
# Print a summary of the coordinate data in $xyz
pdb$xyz
```


```{r}
# Examine the row and column dimensions
dim(pdb$xyz)
```

```{r}
# Print coordinates for the first two atom
pdb$xyz[ 1, atom2xyz(1:2) ]
```

Side-Note: The 'pdb' class. Objects created by the read.pdb() function are of class "pdb". This is recognized by other so called generic Bio3D functions (for example atom.select(), nma(), print(), summary() etc.). A generic function is a function that examines the class of its first argument, and then decides what type of operation to perform (more specifically it decides which specific method to dispatch to). So for example, the generic atom.select() function knows that the input is of class "pdb", rather than for example an AMBER parameter and topology file, and will act accordingly.


Q8. Use the Bio3D write.pdb() function to write out a protein only PDB file for viewing inVMD. Also write out a second separate PDB file for the ligand with residue name MK1
HINT: In Bio3D you can use the trim.pdb() function together with the pub object and your atom selections such as atom.select(pdb, ’protein’) for the fist file and atom.select(pdb, ‘ligand’) for the second.

Let's select the indices of the residues/atoms corresponding to either the protein, or the ligand.
```{r}
inds.protein <- atom.select(pdb, "protein") #retrieve indices of the atoms corresponding to the term protein
inds.protein #note that we have three xyz values per atom
inds.ligand <- atom.select(pdb, "ligand")
inds.ligand
```

Let's see if we can select the correct atoms from the whole pdb structure
```{r}
pdb$atom # the whole thing
```


```{r}
pdb$atom[inds.protein$atom,]  #atoms from pdb that correspond to protein
pdb$atom[inds.ligand$atom,] #atoms from pdb that correspond to ligand
```


```{r}
pdb.ligand <- trim.pdb(pdb, inds=inds.ligand) #trim the pdb for ligand
pdb.ligand
pdb.protein <- trim.pdb(pdb, inds = inds.protein)
pdb.protein
```

```{r}
write.pdb(pdb.ligand, file = "1hsg_ligand.pdb")#write the corresponding .pdb files
write.pdb(pdb.protein, file = "1hsg_protein.pdb")
```

#Section 4: Working with multiple PDB files

```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```
```{r}
pdbs <- pdbaln(files, fit=TRUE)
```

```{r}
# Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)
# Print to screen a summary of the 'pdbs' object
pdbs
# Access the first 5 rows, and 8 columns
pdbs$ali[1:5, 1:8]
# Associated residues numbers
pdbs$resno[1:5, 1:8]
```



##Basic structure analysis
```{r}
# Calculate sequence identity
seqidentity(pdbs)
# Calculate RMSD
rmsd(pdbs)
# Calculate RMSD
rd <- rmsd(pdbs)
# Clustering
hc <- hclust(as.dist(rd))
grps <- cutree(hc, k=3)
# Plot results as dendrogram
hclustplot(hc, k=3)



```



#Section 5:  Example Application on Adenylate Kinase (Adk)

```{r}
aa <- get.seq("1ake_A")
# Blast or hmmer search
b <- blast.pdb(aa)
```


```{r}
# Plot a summary of search results
hits <- plot(b)
```


```{r}
head(hits$pdb.id)
```

```{r}
# Fetch PDBs
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip =
TRUE)
# Align structures
pdbs <- pdbaln(files)
# Vector containing PDB codes
ids <- basename.pdb(pdbs$id)
# Draw schematic alignment
plot(pdbs, labels=ids)
```

###5.2 Sequence conservation analysis
```{r}
# Calculate sequence conservation
cons <- conserv(pdbs, method="entropy22")
# SSE annotations
sse <- pdbs2sse(pdbs, ind=1, rm.gaps=FALSE)
# Plot conservation per residue
plotb3(cons, sse=sse, ylab="Sequence entropy")
```


###5.3 Annotate collected PDB structures
Function pdb.annotate() provides a convenient way of annotating the PDB files we have
collected. Below we use the function to annotate each structure to its source species. This will
come in handy when annotating plots later on:

```{r, eval = FALSE}
library(XML)
library(bitops)
library(RCurl)
anno <- pdb.annotate(ids)
print(unique(anno$source))
```

##5.4 Principal component analysis
Function pca() provides principal component analysis (PCA) of the structure data. PCA is a
statistical approach used to transform a data set down to a few important components that
describe the directions where there is most variance. In terms of protein structures PCA is used
to capture major structural variations within an ensemble of structures.
Function pca() provides principal component analysis (PCA) of the structure data. PCA is a
statistical approach used to transform a data set down to a few important components that
describe the directions where there is most variance. In terms of protein structures PCA is used
to capture major structural variations within an ensemble of structures.
PCA can be performed on the structural ensemble (stored in the pdbs object) with function
pca.xyz(). To obtain meaningful results we first superimpose all structures on the invariant
core (function core.find()).

```{r}
# find invariant core
core <- core.find(pdbs)
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)
# Perform PCA
pc.xray <- pca(pdbs)
# Calculate RMSD
rd <- rmsd(pdbs)
# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)
plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```
```{r}
plot(pc.xray)
```


One can then use the identify() function to label and individual points.
```{r,eval = FALSE}
# Left-click on a point to label and right-click to end
identify(pc.xray$z[,1:2], labels=basename.pdb(pdbs$id))
```

To visualize the major structural variations in the ensemble the function mktrj() can be used to
generate a trajectory PDB file by interpolating along a give PC (eigenvector):
```{r}
# Visualize first principal component
mktrj(pc.xray, pc=1, file="pc_1.pdb")
```


##5.5 Plotting results with ggplot2

```{r}
library(ggplot2)
library(ggrepel)
df <- data.frame(x=pc.xray$z[,1], y=pc.xray$z[,2])
col <- as.factor(grps.rd)
p <- ggplot(df, aes(x, y)) +
 geom_point(aes(col=col), size=2) +
 xlab("PC1") +
 ylab("PC2") +
 scale_color_discrete(name="Clusters") +
 geom_text_repel(aes(label=ids))
p
```

