---
title: "Class8_2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Principal Component Analysis (PCA)

Let's retrieve the data and plot it
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV", row.names=1)
mydata$wt1
plot(mydata$wt1)
plot(mydata$ko1)
plot(mydata)
```

Let's call prcomp() to do PCA. However, prcomp() expects the samples to be rows and genes to be columns so we need to first transpose the matrix with the t() function.
```{r}
pca <- prcomp(t(mydata), scale=TRUE) #
attributes(pca) #pca$x contains the PCs for drawing our first graph.
```

look at pca$x
```{r}
pca$x
```

Lets plot the new PCA data: PC1 vs PC2
```{r}
plot(pca$x[,1], pca$x[,2])
```

Calculate the variance
```{r}
 ## Variance captured per PC
pca.var <- pca$sdev^2
pca.var
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) #percent value, rounded to make it look good
pca.var.per
```

Plot the variance
```{r}
barplot(pca.var.per, main="Scree Plot", xlab="Principal Component", ylab="Percent Variation")
```

Make the plot more useful
```{r}
# I want to make a vector of colors for wt and ko samples
colnames(mydata) # I want my factors to be WT and KO only, not WT1 to WT5 and KO1 to KO5
substr( colnames(mydata), 1, 2) #use substring substr() to remove the extra info.
colvec <- as.factor( substr( colnames(mydata), 1, 2) ) #make a vector for colors.
```


Use the new vector to color WT and KO with two different colors
```{r}
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16, 
     xlab=paste0("PC1 (", pca.var.per[1], "%)"), #paste() concatenates strings
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))
text(pca$x[,1], pca$x[,2], labels = colnames(mydata), pos=1) #adding text labels.
#identify(pca$x[,1], pca$x[,2], labels = colnames(mydata)) #run this then click on the points on the graph that you want to identify, press ESC and it will tell you which points those are.
```



