---
title: "Class7"
date: "4/25/2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions again
We can source any file of R code with the 'source()' function.

```{r}
source("http://tinyurl.com/rescale-R")
```


Let's make sure things are here

```{r}
ls()
```



Check our 'rescale()' function is working

```{r}
rescale(0:10)
rescale(c(0:10,NA))
```


use {r,eval=FALSE} to knit through the error

```{r,eval=FALSE}
rescale( c(1:10,"word"))
```


###Revisit 'rescale()', with warning() and stop(): rescale2
  
  Add an if() condition to check if x is numeric with 'is.numeric()'.
```{r}
is.numeric(c(1:10,NA))
is.numeric(c(1:10,"word"))
!is.numeric(1:10)
```
 

test rescale2()
```{r}
rescale2(c(0:10,NA))
```
 
```{r, eval=FALSE}
rescale2(c(0:10,"word"))
```
 
 
 ###both_na() function
 
```{r, eval=FALSE}
x <- c(1:3, NA, NA)
y <- c(3, 4, 5, NA, 2)
both_na(x, y)

y2 <-  c(3, 4, 5, NA, 2, NA)
both_na(x, y2)
```
 
```{r}
x <- c(1:3, NA, NA)
y <- c(3, 4, 5, NA, 2)
both_na3(x,y)
```
 
 
###gene_intersect()

```{r}
x <- df1$IDs
y <- df2$IDs
x
y
intersect(x,y)
x %in% y
```


We can use the logical output in '%in%' to get at our matching data

```{r}
x[ x %in% y]
y[ y %in% x]
```

Let's put these together as columns of a matrix
```{r}
cbind(x[x %in% y], y[y %in% x])

cbind(c("bla","blou"), c("cra","crou"))

rbind(c("bla","blou"), c("cra","crou"))
```

Use prewritten gene_intersect() on x and y

```{r}
gene_intersect(x,y)
```


Use gene_intersect2() this time, which returns the expression data as well.
```{r}
gene_intersect2(df1,df2)
```


Make the output look better with gene_intersect3()
```{r}
gene_intersect3(df1,df2)
```


Make the code look clearer with gene_intersect4() (make it obviously correct)

Use the 'merge()' function
```{r}
?merge
merge(df1, df2)

```



