---
title: "map_colors_function"
author: Long Vo ngoc
date: 04/20/2018
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# header 1
## header 2
### Let's look at the poor old function
This is the function code from Barry's student.

```{r}
#old function
map.colors <- function (value,high.low,palette) {
  proportion <- ((value-high.low[1])/(high.low[2]-high.low[1]))
  index <- round ((length(palette)-1)*proportion)+1
  return (palette[index])
}
```

### Now let's try to understand this function...

```{r}
map.colors2 <- function (x, 
                         high.low = range(x), #default value for high.low
                         palette = cm.colors(100)) {  #default value for palette
  
  # Determine where in the 'high.low' range or values of 'x' lies. 
  percent <- ( (x - high.low[1]) / (high.low[2] - high.low[1]) )
  
  # Where in the 'palette' vector of colors is this 'percent'
  index <- round ( (length(palette) - 1)* percent) + 1
  
  return (palette[index])
}
```

## Working with our own functions

My first function

```{r}
add <- function(x, y=1) {
# Sum the input x and y
  return(x + y)
}
```

Testing my first function
```{r}
add(1:20, 3)
```


Second function: rescale
```{r}
rescale <- function(x,na.rm = TRUE) {
   rng <-range(x)
   (x - rng[1]) / (rng[2] - rng[1])
}
```


Testing 'rescale'
```{r}
rescale(c(1,2,3,NA,5))
```

Update second function: rescale2
```{r}
rescale2 <- function(x,
                    na.rm = TRUE,
                    plot = FALSE) {
   if(na.rm) {
     rng <- range(x, na.rm=na.rm)  #na.rm is a logical that indicates if NA's should be ommited
   } else {
     rng <- range(x)
   }
   
   print("Hello")
   
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   

   
   print("is it me you are looking for?")
   
   if(plot) {
     plot(answer, typ="b", lwd=4)
}
  
   print("I can see it in ...")

   return(answer)
}
```

Testing rescale2
```{r}
rescale2(c(1,2,NA,3,NA,4,5,6), plot = TRUE)
```


# Section 1B
```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```



Breaking it down to better understand the program
```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s1$remark
```

```{r}
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.chainA$helix
```

```{r}
s1.b <- s1.chainA$atom$b
s1.b
```

```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
```



Generalize as a function
```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```






